---
title: "Lab 04 - Visualizing spatial data"
author: "Jessie"
date: "01/31"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}

library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")

library(ggplot2)
library(dplyr)

c <- dennys
d <- laquinta

```

### Exercise 1
What are the dimensions of the Denny’s dataset? 1643 rows 6 columns, the variables are address, city, state, zip, longitude, latitude
```{r q1, message = FALSE}
nrow(dennys)
ncol(dennys)
dim(dennys)
glimpse(dennys)
```

### Exercise 2
909 rows 6 columns, the variables are address, city, state, zip, longitude, latitude

```{r q2}
nrow(laquinta)
ncol(laquinta)
dim(laquinta)
glimpse(laquinta)
```

### Exercise 3
Take a look at the websites that the data come from (linked above). Are there any La Quinta’s locations outside of the US? If so, which countries? What about Denny’s?

la quinta also locates in Canada, China, Arab, new Zealand, Mexico, turkey, chile, Columbia 
Denny's locates in the united states only.
...

### Exercise 4

Now take a look at the data. What would be some ways of determining whether or not either establishment has any locations outside the US using just the data (and not the websites). Don’t worry about whether you know how to implement this, just brainstorm some ideas. Write down at least one as your answer, but you’re welcomed to write down a few options too.

I think I might filter out the zip code that does not belong to the States. For instance, I would filter out the zip codes that are not composed of 5 digits, and also those which does not start with states' zip codes.

### Exercise 5

no rows are found, indicating there is no denny is located outside of the states.

```{r q5}
c %>%
  filter(!(state %in% states$abbreviation))
```

### Exercise 6

...
```{r 6}
c %>%
  mutate(country = "United States")
```

 
### Exercise 7

cancun Mexico
partido lglesias Mexico
Apodaca Mexico
Medellin Colombia 
Monterrey Mexico
Monterrey Mexico
Oshawa Canada
Poza Rica Mexico
San Jose Chiapa Mexico
Puebla mexico
San Luis Potosi Mexico
Tegucigalpa Hunduras

```{r 7}
d %>%
  filter(!(state %in% states$abbreviation))
```
### Exercise 8

...
```{r 8}
library(dplyr)
library(tidyverse)
d %>%
  mutate(country = case_when(state %in% state.abb     ~ "United States",
                             state == "ANT" ~ "Colombia",
                             state %in% c("ON", "BC") ~ "canada",
                             state %in% c("QR", "NL", "PU", "AG", "CH", "FM") ~"Mexico",
                             state == "FM" ~ "Hunduras" )) %>%
  filter(country == "United States")
```
Dennys: DE has the lowest amount of dennys and CA has the highest.
La Quintas: TX has the highest amount of laquintas and foreign countries has the lowest amount. not surprising because it's where people want to spend their holidays.
```{r q9}
d %>%
  count(state)
c  %>%
  count(state)

c %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```


```{r Q10}
c <- c %>%
  mutate(establishment = "Denny's")
d <- d %>%
  mutate(establishment = "La Quinta")
c_d <- bind_rows(c, d)
ggplot(c_d, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment)) +
  geom_point()
```
```{r q11}
c_d %>%
  filter(state == "NC") %>%
ggplot(c_d, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment, alpha=0.4)) +
                     theme_linedraw() +
                  labs (x = "longitude", y = "latitude", title = "NC's dennys and laquinta", fill = "establishment") + 
  geom_point()
#not quite true in NC.
```
```{r q12}
c_d %>%
  filter(state == "TX") %>%
ggplot(c_d, mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment, alpha=0.2)) +
                     theme_linedraw() +
                  labs (x = "longitude", y = "latitude", title = "TX's dennys and laquinta", fill = "establishment") + 
  geom_point()
# quite true in TX 
```








